---
title: "lab_02"
author: "derek willis"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1** Create a new column that divides the expense amounts into the following categories - \$0-\$500, \$500-\$5000, \$5001-\$50000, \$50001\_\$100000, \$100001-\$500000 - giving each an appropriate text value, and count how many expenses are in each group. How would you describe the distribution of expenses in a sentence? Would you change the categories?

**A1.** the distribution is really uneven. Many people/organizations paid in the mid-range which is a significant amount of money. I think I would want to know the range in which individuals are paying vs organizations. 

```{r}
maryland_expenses_date_with_ranges %>%  
  mutate(
  range = case_when(
    (amount >= 0 & amount <= 500) ~ "range 1",
    (amount >= 501 & amount <= 5000) ~ "range 2",
    (amount >= 5001 & amount <= 50000) ~ "range 3",
    (amount >= 50001 & amount <= 100000) ~ "range 4",
    (amount >= 100001 & amount <= 500000) ~ "range 5",
      TRUE ~ as.character(NA)  )
      ) %>%
  group_by(range) %>%
  summarize(total = sum(amount))
```

```{r}
maryland_expenses_date %>%
  group_by(range) %>%
  summarize(total = sum(amount))
```


```{r}

 
```


**Q2.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? 
**A2.** 26 department, the Atmospheric and oceanic science department offered the most

```{r}
umd_courses %>% filter(str_detect(title, "Climate"))  %>% 
  filter(seats >= 1)  %>%  group_by(department) %>% summarise(
    count_classes = n(), total_seats = sum(seats),) %>% arrange(desc(total_seats))
 
```

**Q3.** Of the departments listed in A2, which one is the most interesting to you, and why?
**A3.** Germanic studies, because I wouldn't think the germanic language language had anything to do with climate.

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.
**A4.**   Moore's committee spent the most money toward the very end of his campaign, closer to the election

```{r}
maryland_expenses_date <- maryland_expenses %>% mutate(new_expenditure_date=mdy(expenditure_date))
```
```{r}

```


```{r}
maryland_expenses_date %>%
  mutate(month = floor_date(new_expenditure_date, "month")) %>%
  group_by(month, committee_name) %>%  
  filter(committee_name == "Moore  Wes For Maryland") %>% 
  summarise(total_amount = sum(amount)) %>%
 arrange(desc(total_amount))
```

**Q5.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.
**A5.** How many Departments offer courses discussing racism, colonization and discrimination with at least one seat? Anything note-worthy? Nine departments, 

```{r}
umd_courses %>% filter(str_detect(title, "Racism")  |
                       str_detect(title, "Colonization") |
                      str_detect(title, "Discrimination"))  %>% 
  filter(seats >= 1)  %>%  group_by(department) %>% summarise(
    count_classes = n(), total_seats = sum(seats),) %>% arrange(desc(total_seats))
```
